---
title: "Reproducible documents"
author: "Simon Chang"
format: html
editor: visual
---

```{r setup}
#| message: false
#| warning: false


library(tidyverse)
post_meal_data <- read_csv2(here::here("data/post-meal-insulin.csv"))
small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)
tidier_post_meal_data <- small_post_meal_data |>
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  ) |>
  filter(glykemi == 0) |>
  select(-glykemi)
```


## Showing the data

```{r}
post_meal_data
glimpse(post_meal_data)
```

```{r ggplot}
library(ggplot2)
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```
## Basic histogram

```{r}
#| fig-cap: "Dist. of BMI"
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```
This is my figure (@fig-bmi-histo)

## Excersice: discrete plots
```{r discrete plots}
#| fig-cap: "Grouping"
#| label: fig-group-bar
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```

See @fig-group-bar above for a cool plot!

## Plotting two discrete

```{r two discrete}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```

## Side by side plots

```{r histogram}
post_meal_data |>
  ggplot(aes(x = BMI)) +
  geom_histogram()
```
```{r}
post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```
```{r side-by-side-plots}
#| label: fig-bmi-glycemia
#| fig-cap: "BMI and glycemia, side by side."
#| fig-subcap:
#|   - "Distribution of BMI."
#|   - "Number of those with glycemia."
#|   - "Grouping in relation to glycemia"
#| layout-ncol: 3
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()

ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()

post_meal_data |>
  ggplot(aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```
## Plotting two continuous variables

```{r two continouus}
#| fig-cap: "Scatterplot of BKI against PG"
#| label: fig-bmi-aucpg

ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point(aes(fill = Group)) +
  geom_smooth()
```
## Pushing changes to remote

Oh oh - how exciting!

## Pulling file from Github

Even more exiting

## Selecting Data

```{r}
select(post_meal_data, Age)
select(post_meal_data, Age, BMI, Weight)
select(post_meal_data, -Group)
select(post_meal_data, starts_with("Insulin"))
select(post_meal_data, ends_with("5"))
select(post_meal_data, contains("0"))
```

## Renaming specific columns

```{r rename}
small_post_meal_data |>
  rename(auc_glucose = auc_pg)
```
## Select and rename tasks


```{r}
# Task 1.
post_meal_data |>
  select(contains("OGTT"))

# Task 2.
post_meal_data |>
  select(contains(".."), -contains("...")) |>
  rename(P.Glucose.minus5.OGTT = P.Glucose..5.OGTT) |>
  rename(Insulin.minus5.OGTT = Insulin..5.OGTT.X)


# Task 3.
basic_info <- select(post_meal_data, id, BMI, Length, Weight, Age)
rename(basic_info, Height = Length)
```


## Filtering on rows

```{r}
small_post_meal_data |>
  filter(glykemi == 1)

small_post_meal_data |>
  filter(BMI >= 25)

small_post_meal_data |>
  filter(Group == "CTR")
```
```{r}
TRUE & TRUE
TRUE & FALSE
FALSE & FALSE

TRUE | TRUE
TRUE | FALSE
FALSE | FALSE
```
```{r}
small_post_meal_data |>
  filter(Group == "CTR", BMI >= 25)

small_post_meal_data |>
  filter(Group == "CTR" | BMI >= 25)
```

## Modify or add column

```{r}
tidier_post_meal_data %>%
  mutate(
    insulin_glucose_ratio = auc_insulin / auc_glucose,
    Group = if_else(Group == "CTR", "Conrol", "First-degree Relative")
  )
```
